---
title: "The Behavioral Data Files — Episode 1"
subtitle: "What Skills Actually Get You Hired in AI Right Now?"
author: "Edouard Tamba | GritGateway"
date: "May 2026"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
# Global settings for every chunk in this document.
# echo=FALSE   — hides my code from the final blog output
# message=FALSE — hides package loading messages
# warning=FALSE — hides warnings
# fig settings  — default chart dimensions
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 10,
  fig.height = 6
)
```

```{r packages, message=FALSE, warning=FALSE}
# Loading the four libraries I need for this analysis.
# data.table — loads large files (1.3M rows) efficiently
# tidyverse  — data cleaning + ggplot2 charts
# janitor    — standardises messy column names
# scales     — formats percentages and numbers on charts
library(data.table)
library(tidyverse)
library(janitor)
library(scales)
library(ggtext)

cat("All packages loaded.\n")
```

---

# The Question

Every week thousands of people ask the same thing:

> *"What do I actually need to learn to get hired in AI right now?"*

Job boards give you thousands of postings. Influencers give you opinions.
I gave it **1.3 million LinkedIn job postings** and ran the numbers.

This is what the data actually says.

---

# Step 1 — Peek At The Data

```{r peek-data}
# Before loading 1.3 million rows I peek at 100 rows first.
# This confirms column names and structure without waiting.
# Same logic as: SELECT TOP 100 * FROM table in SQL.
peek_skills <- fread(
  "C:/Users/Aedou/Music/DATA_ANALYSIS/behavioral-data-files/data/raw/job_skills.csv",
  nrows = 100)

peek_postings <- fread(
  "C:/Users/Aedou/Music/DATA_ANALYSIS/behavioral-data-files/data/raw/linkedin_job_postings.csv",
  nrows = 100)

salary <- fread(
  "C:/Users/Aedou/Music/DATA_ANALYSIS/behavioral-data-files/data/raw/ai_job_dataset.csv")

cat("=== JOB SKILLS COLUMNS ===\n");   print(names(peek_skills))
cat("\n=== JOB POSTINGS COLUMNS ===\n"); print(names(peek_postings))
cat("\n=== SALARY COLUMNS ===\n");       print(names(salary))
```

---

# Step 2 — Load Full Datasets

```{r load-full, cache=TRUE}
# Now that I confirmed the structure I load the full files.
# fread() from data.table is built for large files.
# It reads 1.3M rows significantly faster than read_csv().
# cache=TRUE means R saves the result — if I re-knit
# without changing this chunk it skips the 60 second wait.
cat("Loading full datasets — may take 60 seconds...\n")

skills_full <- fread(
  "C:/Users/Aedou/Music/DATA_ANALYSIS/behavioral-data-files/data/raw/job_skills.csv")

postings_full <- fread(
  "C:/Users/Aedou/Music/DATA_ANALYSIS/behavioral-data-files/data/raw/linkedin_job_postings.csv")

cat("Skills rows loaded:   ", nrow(skills_full), "\n")
cat("Postings rows loaded: ", nrow(postings_full), "\n")
cat("Done.\n")
```

---

# Step 3 — Filter AI Roles Only

```{r filter-ai}
# I'm building a search pattern that works like a SQL WHERE
# clause with multiple OR conditions on the job_title column.
# paste() + collapse="|" joins keywords into one regex pattern.
ai_pattern <- paste(c(
  "artificial intelligence", "machine learning",
  "data scientist", "ml engineer", "ai engineer",
  "deep learning", "nlp", "computer vision",
  "data analyst", "ai researcher", "llm",
  "generative ai", "prompt engineer"
), collapse = "|")

# I found non-AI jobs in the first pass — warehouse and
# logistics roles that LinkedIn's algorithm included when
# the data was originally scraped. This exclusion pattern
# removes them. Real world data is never clean.
exclude_pattern <- paste(c(
  "fulfillment", "warehouse", "delivery",
  "driver", "cashier", "retail", "forklift",
  "associate", "team lead", "package"
), collapse = "|")

# STEP 1: Keep rows where job_title matches an AI keyword.
# tolower() makes matching case-insensitive.
# SQL: WHERE LOWER(job_title) LIKE '%data scientist%' OR ...
ai_postings <- postings_full %>%
  filter(str_detect(tolower(job_title), ai_pattern))

# STEP 2: Remove false positives.
# ! means NOT — keep everything that does NOT match exclusions.
# SQL: AND LOWER(job_title) NOT LIKE '%fulfillment%' AND ...
ai_postings_clean <- ai_postings %>%
  filter(!str_detect(tolower(job_title), exclude_pattern))

cat("Total postings in dataset: ", nrow(postings_full), "\n")
cat("AI roles before cleaning:  ", nrow(ai_postings), "\n")
cat("False positives removed:   ",
    nrow(ai_postings) - nrow(ai_postings_clean), "\n")
cat("AI roles after cleaning:   ", nrow(ai_postings_clean), "\n")
cat("% of total that are AI:    ",
    round(nrow(ai_postings_clean)/nrow(postings_full)*100, 1), "%\n")

cat("\nTop 10 AI job titles after cleaning:\n")
print(ai_postings_clean %>%
        count(job_title, sort = TRUE) %>%
        slice_head(n = 10))
```

---

# Step 4 — Join Skills to AI Roles

```{r join-skills}
# I join the skills file to my clean AI postings using
# job_link as the shared key between both tables.
#
# SQL equivalent:
# SELECT s.job_skills, p.job_title, p.company, p.job_location
# FROM skills_full s
# INNER JOIN ai_postings_clean p ON s.job_link = p.job_link
#
# INNER JOIN = only rows where job_link exists in BOTH tables.
# Jobs with no skills listed are dropped.
# Skills with no matching AI job are dropped.
ai_skills <- skills_full %>%
  inner_join(ai_postings_clean, by = "job_link") %>%
  clean_names()

cat("Rows after join:     ", nrow(ai_skills), "\n")
cat("Unique AI jobs:      ", n_distinct(ai_skills$job_link), "\n")
cat("Unique skills found: ", n_distinct(ai_skills$job_skills), "\n")
```

---

# Step 5 — Explode Skills to One Per Row

```{r explode-skills}
# My job_skills column has all skills in one cell separated
# by commas — like a list inside a single database field.
# I need to EXPLODE it so each skill gets its own row.
# This is going from WIDE to LONG format.
# PostgreSQL equivalent: unnest(string_to_array(job_skills,','))

ai_skills_long <- ai_skills %>%
  separate_rows(job_skills, sep = ",") %>%

  # str_trim removes whitespace around skill names
  # so "Python" not " Python"
  mutate(job_skills = str_trim(job_skills)) %>%

  # Remove empty rows created during splitting
  filter(job_skills != "" & !is.na(job_skills))

cat("Rows BEFORE exploding:", nrow(ai_skills), "\n")
cat("Rows AFTER exploding: ", nrow(ai_skills_long), "\n")
cat("Average skills per job:",
    round(nrow(ai_skills_long)/n_distinct(ai_skills_long$job_link), 0), "\n")
```

---

# Step 6 — Standardise & Count Top Skills

```{r standardise-count}
# PROBLEM 1: Same skills counted twice due to capitalisation
# "Machine Learning" and "Machine learning" = two counts
# FIX: str_to_title() standardises everything to Title Case
# so all variations merge into one before counting.
#
# PROBLEM 2: Acronyms broken by Title Case
# "SQL" becomes "Sql", "AWS" becomes "Aws"
# FIX: case_when() manually corrects known acronyms.
# This is R's version of SQL CASE WHEN:
# CASE WHEN job_skills = 'Sql' THEN 'SQL' ELSE job_skills END
#
# PROBLEM 3: "Communication" and "Communication Skills"
# are the same skill with two different labels.
# FIX: both map to "Communication" in case_when.

ai_skills_clean <- ai_skills_long %>%
  mutate(
    # Step 1 — Standardise to Title Case first
    job_skills = str_to_title(str_trim(job_skills)),

    # Step 2 — Fix acronyms and merge duplicate labels.
    # TRUE ~ job_skills is the ELSE clause —
    # anything not listed here keeps its current name.
    job_skills = case_when(
      job_skills == "Sql"                  ~ "SQL",
      job_skills == "Aws"                  ~ "AWS",
      job_skills == "Pytorch"              ~ "PyTorch",
      job_skills == "Tensorflow"           ~ "TensorFlow",
      job_skills == "Power Bi"             ~ "Power BI",
      job_skills == "Nlp"                  ~ "NLP",
      job_skills == "Llm"                  ~ "LLM",
      job_skills == "Communication Skills" ~ "Communication",
      job_skills == "R Programming"        ~ "R",
      TRUE                                 ~ job_skills
    )
  )

# I tag soft skills separately so the chart tells a richer story —
# technical vs human skills both matter in AI roles.
# This becomes a key insight in the LinkedIn post.
soft_skills <- c(
  "Communication", "Collaboration", "Teamwork",
  "Problem Solving", "Project Management",
  "Leadership", "Critical Thinking",
  "Attention To Detail", "Time Management",
  "Presentation Skills"
)

# Count unique jobs requiring each skill after standardisation.
# I use n_distinct(job_link) not n() to count jobs, not rows.
# SQL equivalent:
# SELECT job_skills, COUNT(DISTINCT job_link) as job_count
# FROM ai_skills_clean
# GROUP BY job_skills
# ORDER BY job_count DESC
# LIMIT 25
top_skills <- ai_skills_clean %>%
  group_by(job_skills) %>%
  summarise(
    job_count = n_distinct(job_link),
    .groups = "drop"
  ) %>%
  arrange(desc(job_count)) %>%
  slice_head(n = 25) %>%
  mutate(
    # Calculate what % of all AI jobs require this skill
    pct = round(
      job_count / n_distinct(ai_skills_clean$job_link) * 100, 1),

    # Tag each skill — drives colour coding in the chart
    skill_type = ifelse(
      job_skills %in% soft_skills, "Soft Skill", "Technical Skill"),

    # Reorder factor so highest bar appears at top of chart
    job_skills = fct_reorder(job_skills, pct)
  )

cat("=== TOP 25 AI SKILLS AFTER STANDARDISATION ===\n")
print(top_skills %>%
        select(job_skills, job_count, pct, skill_type), n = 25)
```

---

# Step 7 — The Chart
```{r}
#install.packages("ggtext")
library(ggtext)
# ==============================================================
# STORYTELLING WITH DATA — KNAFLIC PRINCIPLES APPLIED
# VERSION 3 — All visual issues resolved
# ==============================================================
#
# PRINCIPLE 1 — RIGHT CHART TYPE
# Horizontal bars for long labels — readers scan naturally
# top to bottom matching our ranked order.
#
# PRINCIPLE 2 — ACTION TITLE
# Title communicates the TAKEAWAY not just the topic.
#
# PRINCIPLE 3 — COLOUR WITH PURPOSE
# Blue  = top 4 core non-negotiables
# Amber = soft skills (the surprise finding)
# Grey  = all other technical skills (recede)
#
# PRINCIPLE 4 — LABEL DATA DIRECTLY
# Percentage on every bar — no axis needed.
#
# PRINCIPLE 5 — REMOVE ALL CLUTTER
# No border, no gridlines, no legend, no axis, no ticks.
# Rect annotation removed — it was hijacking the Teamwork label.
# Knaflic: if an element causes confusion, remove it.
#
# PRINCIPLE 6 — ANNOTATE KEY INSIGHTS
# Primary annotation on Communication — the big surprise.
# Secondary annotation at bottom — confirms soft skills
# are a pattern across top 20, not just one fluke.
# ==============================================================

# Trim to top 20 — remove noise below 11%
top_20 <- top_skills %>%
  slice_head(n = 20) %>%
  mutate(
    bar_color = case_when(
      job_skills %in% c("Python", "Machine Learning",
                        "SQL", "Data Analysis") ~ "Top Technical",
      skill_type == "Soft Skill"                ~ "Soft Skill",
      TRUE                                      ~ "Other Technical"
    )
  )

ggplot(top_20,
       aes(x    = job_skills,
           y    = pct,
           fill = bar_color)) +

  # BARS
  geom_col(width = 0.65) +

  # DIRECT LABELS — bold on highlighted bars, plain on grey
  # Knaflic: label data directly, remove axis entirely
  geom_text(
    aes(
      label    = paste0(pct, "%"),
      fontface = ifelse(bar_color == "Other Technical",
                        "plain", "bold")
    ),
    hjust = -0.12,
    size  = 3.7,
    color = "#374151"
  ) +

  # PRIMARY ANNOTATION — Communication is the surprise
  # Knaflic: direct the reader to the one thing that matters
  annotate(
    geom       = "text",
    x          = "Communication",
    y          = 54,
    label      = "Beats R, Tableau\n& TensorFlow",
    hjust      = 0,
    size       = 3.4,
    color      = "#92400e",
    fontface   = "bold.italic",
    lineheight = 1.3
  ) +

  # Arrow — yend=42 keeps tip clear of the 37.1% label
  # Missing comma after yend was causing syntax error — fixed
  annotate(
    geom      = "segment",
    x         = "Communication",
    xend      = "Communication",
    y         = 53,
    yend      = 42,
    color     = "#92400e",
    linewidth = 0.6,
    arrow     = arrow(length = unit(0.18, "cm"),
                      type   = "closed")
  ) +

  # SECONDARY ANNOTATION — placed at Power BI (bottom bar)
  # Moved away from Teamwork which was hijacking its label
  # Knaflic: annotation should clarify, never create confusion
  annotate(
    geom       = "text",
    x          = "Power BI",
    y          = 30,
    label      = "5 of the top 20 skills\nare soft skills",
    hjust      = 0,
    size       = 3.1,
    color      = "#92400e",
    fontface   = "italic",
    lineheight = 1.3
  ) +

  coord_flip() +

  # COLOURS — purposeful not decorative
  scale_fill_manual(
    values = c(
      "Top Technical"   = "#2563EB",
      "Soft Skill"      = "#f59e0b",
      "Other Technical" = "#d1d5db"
    )
  ) +

  # Remove x-axis — direct labels replace it
  # Knaflic: remove axis when data is labelled directly
  scale_y_continuous(
    limits = c(0, max(top_20$pct) * 1.42),
    labels = NULL,
    breaks = NULL
  ) +

  labs(
    title = "Python Is Essential But Soft Skills Beat Most AI Tools",

    # ggtext HTML subtitle — actual coloured squares
    # instead of plain text saying colour names
    subtitle = "% of 4,400+ AI job postings on LinkedIn (2024) requiring each skill<br>
<span style='color:#2563EB'>■ Core non-negotiables</span> &nbsp;&nbsp;
<span style='color:#f59e0b'>■ Soft skills</span> &nbsp;&nbsp;
<span style='color:#b0b7c3'>■ Other technical skills</span>",

    caption = "Source: LinkedIn Job Postings & Skills Dataset · 1.3M rows · Kaggle 2024\nAnalysis: Edouard AGBOR | The Behavioral Data Files — Episode 1 | gritgateway.com",
    x = NULL,
    y = NULL
  ) +

  theme_minimal(base_size = 12) +
  theme(

    # Bold dark title — first thing the eye lands on
    plot.title = element_text(
      face       = "bold",
      size       = 17,
      color      = "#0f172a",
      lineheight = 1.2,
      margin     = margin(b = 4)
    ),

    # element_markdown activates HTML colour squares in subtitle
    plot.subtitle = element_markdown(
      color      = "#64748b",
      size       = 10.5,
      lineheight = 1.6,
      margin     = margin(b = 16)
    ),

    # Source — very muted, left aligned
    # Knaflic: left align for cleaner visual lines
    plot.caption = element_text(
      color      = "#9ca3af",
      size       = 8.5,
      lineheight = 1.4,
      hjust      = 0
    ),

    # Skill names — readable, not competing with bars
    axis.text.y = element_text(
      color  = "#1e293b",
      size   = 11.5,
      margin = margin(r = 4)
    ),

    # Knaflic declutter checklist — all removed
    legend.position  = "none",
    panel.grid       = element_blank(),
    panel.border     = element_blank(),
    axis.ticks       = element_blank(),
    plot.background  = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin      = margin(25, 45, 20, 20)
  )

# Full absolute path — consistent with all fread() calls
ggsave(
  filename = "C:/Users/Aedou/Music/DATA_ANALYSIS/behavioral-data-files/outputs/plots/01_top_ai_skills.png",
  width    = 13,
  height   = 10,
  dpi      = 300,
  bg       = "white"
)

cat("Chart saved.\n")

```


```{r}
# Loading the REAL salary data from aijobs.net (now foorilla.com)
# Source: anonymous professional salary submissions worldwide
# License: CC0 public domain — free to use and publish
# This replaces the synthetic Kaggle dataset entirely.

salary_real <- fread(
  "C:/Users/Aedou/Music/DATA_ANALYSIS/behavioral-data-files/data/raw/salaries_real.csv"
)

# Integrity check — confirm it loaded correctly
cat("=== REAL SALARY DATA INTEGRITY CHECK ===\n")
cat("Rows:   ", nrow(salary_real), "\n")
cat("Cols:   ", ncol(salary_real), "\n")

cat("\nYears covered:\n")
print(table(salary_real$work_year))

cat("\nExperience level distribution:\n")
print(table(salary_real$experience_level))

cat("\nEmployment type distribution:\n")
print(table(salary_real$employment_type))

cat("\nSalary range (USD):\n")
cat("Min:    $", format(min(salary_real$salary_in_usd,
                            na.rm=TRUE), big.mark=","), "\n")
cat("Max:    $", format(max(salary_real$salary_in_usd,
                            na.rm=TRUE), big.mark=","), "\n")
cat("Median: $", format(median(salary_real$salary_in_usd,
                               na.rm=TRUE), big.mark=","), "\n")
cat("Mean:   $", format(round(mean(salary_real$salary_in_usd,
                               na.rm=TRUE)), big.mark=","), "\n")

cat("\nTop 15 job titles by count:\n")
print(
  salary_real %>%
    count(job_title, sort = TRUE) %>%
    slice_head(n = 15)
)

cat("\nMissing values per column:\n")
print(colSums(is.na(salary_real)))
```

```{r}
# I'm using REAL salary data from aijobs.net — anonymous
# professional submissions, not synthetic or modelled data.
# 151,445 rows. Zero missing values. Real distributions.
#
# STRATEGY: No skills column exists in salary data.
# I use job title as a PROXY for skills.
# ML Engineer → Python, TensorFlow, PyTorch
# Data Analyst → SQL, Tableau, Excel
# Data Scientist → Python, R, Statistics, SQL
# This is the honest professional approach.
#
# FILTERS APPLIED:
# 1. 2024-2025 only — aligns with LinkedIn skills data period
# 2. Full-time only — 99.4% of data, removes outliers
# 3. AI/data roles only — removes generic "Manager", "Engineer"
# 4. Minimum 100 records per title — statistical reliability

# Define AI/data role keywords
salary_ai_pattern <- paste(c(
  "data scientist", "data analyst", "data engineer",
  "machine learning", "ai engineer", "ml engineer",
  "research scientist", "applied scientist",
  "analytics engineer", "computer vision",
  "nlp engineer", "deep learning", "ai architect",
  "business intelligence", "data architect",
  "mlops", "llm engineer", "prompt engineer"
), collapse = "|")

# STEP 1 — Filter to relevant rows
salary_filtered <- salary_real %>%
  filter(
    work_year    %in% c(2024, 2025),
    employment_type == "FT",
    str_detect(tolower(job_title), salary_ai_pattern)
  )

cat("Rows after filtering:    ", nrow(salary_filtered), "\n")
cat("Unique job titles:       ",
    n_distinct(salary_filtered$job_title), "\n")
cat("Year split:\n")
print(table(salary_filtered$work_year))

# STEP 2 — Median salary by job title
# Median not mean — salary distributions are right-skewed.
# A few $800k outliers would drag the mean up dishonestly.
salary_by_role <- salary_filtered %>%
  group_by(job_title) %>%
  summarise(
    median_salary = median(salary_in_usd, na.rm = TRUE),
    mean_salary   = round(mean(salary_in_usd, na.rm = TRUE)),
    sample_size   = n(),
    .groups = "drop"
  ) %>%
  # Only roles with 100+ data points — statistically reliable
  filter(sample_size >= 100) %>%
  arrange(desc(median_salary)) %>%
  slice_head(n = 15)

cat("\n=== MEDIAN SALARY BY AI/DATA ROLE (top 15) ===\n")
print(salary_by_role %>%
        mutate(
          median_salary = paste0("$",
                          format(median_salary, big.mark = ",")),
          mean_salary   = paste0("$",
                          format(mean_salary,   big.mark = ","))
        ))
```


```{r}
# I'm manually mapping job roles to the skills from our
# top 20 LinkedIn analysis. No skills column exists in
# salary data so this bridge is necessary and transparent.
# This is a judgment call based on industry knowledge —
# I document it so readers can audit my logic.

skill_salary_map <- tribble(
  ~skill,              ~primary_role,                ~median_salary,
  "Python",            "Data Scientist",                     150000,
  "Machine Learning",  "Machine Learning Engineer",          190000,
  "SQL",               "Data Engineer",                      140900,
  "Data Analysis",     "Analytics Engineer",                 145000,
  "Communication",     "Data Scientist",                     150000,
  "Data Visualization","Business Intelligence Engineer",     142050,
  "R",                 "Research Scientist",                 183825,
  "Data Science",      "Data Scientist",                     150000,
  "Tableau",           "Business Intelligence Engineer",     142050,
  "TensorFlow",        "Machine Learning Engineer",          190000,
  "PyTorch",           "Machine Learning Engineer",          190000,
  "Statistics",        "Research Scientist",                 183825,
  "AWS",               "MLOps Engineer",                     160000,
  "Deep Learning",     "Machine Learning Scientist",         165000,
  "Power BI",          "Business Intelligence Engineer",     142050,
  "Project Management","Data Architect",                     151898,
  "Problem Solving",   "Data Scientist",                     150000,
  "Data Analytics",    "Analytics Engineer",                 145000,
  "Computer Vision",   "Computer Vision Engineer",           170200,
  "Data Mining",       "Data Engineer",                      140900
) %>%
  # Tag skill type — same categories as the demand chart
  mutate(
    skill_type = case_when(
      skill %in% c("Communication", "Problem Solving",
                   "Project Management")    ~ "Soft Skill",
      median_salary >= 165000              ~ "High Salary Technical",
      TRUE                                 ~ "Standard Technical"
    ),
    # Format for chart labels
    salary_label = paste0("$",
                   format(median_salary, big.mark=",", trim=TRUE)),
    # Reorder for chart
    skill = fct_reorder(skill, median_salary)
  )

cat("=== SKILL TO SALARY MAPPING ===\n")
print(skill_salary_map %>%
        arrange(desc(median_salary)) %>%
        select(skill, primary_role, median_salary, skill_type))
```



```{r}
# Merge Data Analytics into Data Analysis — same salary, same role.
# Two bars at identical height looks like a data error to readers.
skill_salary_map <- skill_salary_map %>%
  filter(skill != "Data Analytics")

# Salary chart — Knaflic principles
# Reference line at $150k = Data Scientist benchmark
# Blue = premium skills, Amber = soft skills, Grey = standard

ggplot(skill_salary_map,
       aes(x = skill, y = median_salary, fill = skill_type)) +

  # Dashed reference line at $150k benchmark
  geom_hline(yintercept = 150000,
             color      = "#94a3b8",
             linewidth  = 0.5,
             linetype   = "dashed") +

  geom_col(width = 0.65) +

  # Direct salary labels
  geom_text(
    aes(label    = salary_label,
        fontface = ifelse(skill_type == "High Salary Technical",
                          "bold", "plain")),
    hjust = -0.12,
    size  = 3.5,
    color = "#374151") +

  # PRIMARY ANNOTATION — moved higher to clear bar label
  annotate("text",
    x        = "TensorFlow",
    y        = 222000,
    label    = "ML specialisation pays\n$40k more than Python alone",
    hjust    = 0,
    size     = 3.2,
    color    = "#1e40af",
    fontface = "bold.italic",
    lineheight = 1.3) +

  # Arrow — starts below annotation, ends at bar tip
  annotate("segment",
    x         = "TensorFlow",
    xend      = "TensorFlow",
    y         = 219000,
    yend      = 208000,
    color     = "#1e40af",
    linewidth = 0.6,
    arrow     = arrow(length = unit(0.18, "cm"),
                      type   = "closed")) +

  # SECONDARY ANNOTATION — SQL tension
  annotate("text",
    x        = "SQL",
    y        = 175000,
    label    = "#3 most demanded\nbut lowest salary",
    hjust    = 0,
    size     = 3.0,
    color    = "#64748b",
    fontface = "italic",
    lineheight = 1.3) +

  # BENCHMARK LABEL — top left, always visible inside plot
  annotate("text",
    x        = 19,
    y        = 152500,
    label    = "$150k benchmark →",
    hjust    = 1,
    size     = 2.9,
    color    = "white",
    fontface = "italic") +

  coord_flip() +

  scale_fill_manual(values = c(
    "High Salary Technical" = "#2563EB",
    "Standard Technical"    = "#d1d5db",
    "Soft Skill"            = "#f59e0b")) +

  scale_y_continuous(
    limits = c(0, max(skill_salary_map$median_salary) * 1.35),
    labels = NULL,
    breaks = NULL) +

  labs(
    title    = "The Skills That Pay Most Are NOT the Most Demanded",
    subtitle = "Median salary mapped to top 20 AI skills | Real data: aijobs.net · 65k+ submissions · 2024–2025<br>
<span style='color:#2563EB'>■ Premium salary (≥$165k)</span> &nbsp;&nbsp;
<span style='color:#f59e0b'>■ Soft skills</span> &nbsp;&nbsp;
<span style='color:#b0b7c3'>■ Standard technical</span>",
    caption  = "Salary: aijobs.net real survey · CC0 license | Skills demand: LinkedIn 1.3M rows · Kaggle 2024\nAnalysis: Edouard AGBOR | The Behavioral Data Files — Episode 1 | gritgateway.com",
    x = NULL,
    y = NULL) +

  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(
      face = "bold", size = 16,
      color = "#0f172a", lineheight = 1.2,
      margin = margin(b = 4)),
    plot.subtitle = element_markdown(
      color = "#64748b", size = 10,
      lineheight = 1.6, margin = margin(b = 14)),
    plot.caption  = element_text(
      color = "#9ca3af", size = 8,
      lineheight = 1.5, hjust = 0),
    axis.text.y   = element_text(
      color = "#1e293b", size = 11,
      margin = margin(r = 4)),
    legend.position  = "none",
    panel.grid       = element_blank(),
    panel.border     = element_blank(),
    axis.ticks       = element_blank(),
    plot.background  = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    # Increased bottom margin — benchmark label now visible
    plot.margin      = margin(25, 55, 35, 20))

ggsave(
  "C:/Users/Aedou/Music/DATA_ANALYSIS/behavioral-data-files/outputs/plots/02_salary_by_skill.png",
  width = 13, height = 10, dpi = 300, bg = "white")

cat("Salary chart saved.\n")
```


































---

# Step 8 — Save Clean Data

```{r save}
# I save the clean results so I can reuse them later —
# for the blog, for GitHub, and for future episodes.
# This is my contribution to the public data ecosystem.
write_csv(top_skills,
          "data/clean/top_ai_skills_2024.csv")

cat("Clean data saved to data/clean/\n")
```

---

# The Finding

> *We fill this section together once the chart is confirmed.*

---

# About This Analysis

**Data:** LinkedIn Job Postings & Skills — 1.3M rows (Kaggle 2024)

**Tools:** R · data.table · tidyverse · ggplot2

**Author:** Edouard Tamba —
Founder of GritGateway | M.S. Applied Human-Centered AI,
Syracuse University | Chevening Scholar | Mandela Washington Fellow

---

*Next episode: Do people who study abroad return home?*